---
output: github_document
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)  
```

# readabs
<!-- badges: start -->
[![Build Status](https://travis-ci.org/MattCowgill/readabs.svg?branch=master)](https://travis-ci.org/MattCowgill/readabs)
[![codecov status](https://img.shields.io/codecov/c/github/mattcowgill/readabs.svg)](https://codecov.io/gh/MattCowgill/readabs)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![CRAN status](https://www.r-pkg.org/badges/version/readabs)](https://cran.r-project.org/package=readabs)
[![R build status](https://github.com/mattcowgill/readabs/workflows/R-CMD-check/badge.svg)](https://github.com/mattcowgill/readabs/actions)
<!-- badges: end -->

## Overview
{readabs} helps you easily download, import, and tidy time series data from the Australian Bureau of Statistics within R. 
This saves you time manually downloading and tediously tidying time series data and allows you to spend more time on your 
analysis.

Github issues containing error reports or feature requests are welcome. 
Alternatively you can email the package maintainer at mattcowgill at gmail dot com.

## Installation

Install the latest CRAN version of {readabs} with:

```{r cran-installation, eval = FALSE}
install.packages("readabs")
```

You can install the developer version of {readabs} from GitHub with:
```{r gh-installation, eval = FALSE}
# if you don't have devtools installed, first run:
# install.packages("devtools")
devtools::install_github("mattcowgill/readabs")
```

## Usage

There is one key function in {readabs}. It is `read_abs()`, which downloads, imports, and tidies time series data from the ABS website.

There are some other functions you may find useful.

* `read_abs_local()` imports and tidies time series data from ABS spreadsheets stored on a local drive.
* `separate_series()` splits the `series` column of a tidied ABS time series spreadsheet into multiple columns, reducing the manual wrangling that's needed to work with the data.

Both `read_abs()` and `read_abs_local()` return a single tidy data frame (tibble).

## Examples

To download all the time series data from an ABS catalogue number to your disk, and import the data to R as a single tidy data frame, use `read_abs()`. 

First we'll load {readabs} and the {tidyverse}:
```{r load-packages, results=FALSE, warning=FALSE}
library(readabs)
library(tidyverse)
```

Now we'll create one data frame that contains all the time series data from the Wage Price Index, catalogue number 6345.0:

```{r all-wpi}
all_wpi <- read_abs("6345.0")
```

This is what it looks like:

```{r str-wpi}
str(all_wpi)
```

It only takes you a few lines of code to make a graph from your data:

```{r all-in-one-example}
all_wpi %>%
  filter(series == "Percentage Change From Corresponding Quarter of Previous Year ;  Australia ;  Total hourly rates of pay excluding bonuses ;  Private and Public ;  All industries ;",
         !is.na(value)) %>%
  ggplot(aes(x = date, y = value, col = series_type)) +
  geom_line() +
  theme_minimal() +
  labs(y = "Annual wage growth (per cent)")

```


In the example above we downloaded all the time series from a catalogue number. This will often be overkill. If you know the data you need is in a particular table, you can just get that table like this:

```{r wpi1}
wpi_t1 <- read_abs("6345.0", tables = 1)
```

If you want multiple tables, but not the whole catalogue, that's easy too:

```{r wpi1_5}
wpi_t1_t5 <- read_abs("6345.0", tables = c("1", "5a"))
```

In most cases, the `series` column will contain multiple components, separated by ';'. The `separate_series()` function can help wrangling this column.

For more examples, please see the readabs vignette (run `browseVignettes("readabs")`).

## New in recent versions

In 0.4.3,

* `read_abs()` checks to see if you are connected to the internet, and can connect to the ABS website, before trying to download data. The method used for checking connectivity has been improved, to make the function compatible with a broader range of IT environments. Thanks to Oscar Lane for noticing the issue and proposing a fix.

In 0.4.2,

* The default path for files downloaded by `read_abs()` is now set by an environment variable, rather than defaulting to 'data/ABS'. If the variable is not set, downloaded files will be stored in a temporary directory. Thanks to Hugh Parsonage for the idea and implementation. See `?read_abs` for more information.

In 0.4.1,

* The `separate_series()` function gained a `remove_nas` argument - thanks to Sam Gow for the suggestion. This makes tidying data easier.
* Various minor improvements and bug fixes.

In 0.4.0, 

* A new `separate_series()` function was added, to be used in conjunction with `read_abs()` to help tidy data (thanks to David Diviny).
* A new `read_cpi()` convenience function was added; it returns a tibble containing dates and CPI index numbers.
* The `read_abs()` function now has an optional `series_id` argument, allowing users to get individual time series by using the unique identifiers given to them by the ABS.
* The `read_abs()` function now has a `retain_files` argument (`TRUE` by default); by setting this to `FALSE`, downloaded spreadsheets will be stored in a temporary directory
* Some minor bug fixes and enhances - see NEWS for details.

## Awesome Official Statistics Software

[![Mentioned in Awesome Official Statistics ](https://awesome.re/mentioned-badge.svg)](https://github.com/SNStatComp/awesome-official-statistics-software)

We're pleased to be included in a [list of software](https://github.com/SNStatComp/awesome-official-statistics-software) that can be used to work with official statistics.

## Package history

From version 0.3.0, `readabs` gained significant new functionality and the package changed substantially. 
Pre-0.3.0 functions still work, but `read_abs_data()` is deprecated. The behaviour of `read_abs_metadata()` has changed and the function is deprecated. The old version of `readabs` is available in the [0.2.9 branch on Github](https://github.com/MattCowgill/readabs/tree/0.2.9).
