---
output: github_document
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)  
```

# readabs
<!-- badges: start -->
[![Build Status](https://travis-ci.org/MattCowgill/readabs.svg?branch=master)](https://travis-ci.org/MattCowgill/readabs)
[![codecov status](https://img.shields.io/codecov/c/github/mattcowgill/readabs.svg)](https://codecov.io/gh/MattCowgill/readabs)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![packageversion](https://img.shields.io/badge/Package%20version-`r version`-orange.svg?style=flat-square)](commits/master)
[![CRAN status](https://www.r-pkg.org/badges/version/readabs)](https://cran.r-project.org/package=readabs)
<!-- badges: end -->

## Overview
readabs contains tools to easily download, import, and tidy time series data from the Australian Bureau of Statistics. This saves you time manually downloading and
tediously tidying time series data and allows you to spend more time on your 
analysis.

We'd welcome Github issues containing error reports or feature requests. 
Alternatively you can email the package maintainer at mattcowgill at gmail dot com.

## Installation

Install the latest CRAN version of **readabs** with:

```{r cran-installation, eval = FALSE}
install.packages("readabs")
```

You can install the developer version of **readabs** from GitHub with:
```{r gh-installation, eval = FALSE}
# if you don't have devtools installed, first run:
# install.packages("devtools")
devtools::install_github("mattcowgill/readabs")
```

## Usage

There are two key functions in **readabs**. They are:

* `read_abs()` downloads, imports, and tidies time series data from the ABS website.
* `read_abs_local()` imports and tidies time series data from ABS spreadsheets stored on a local drive.

Both functions return a single tidy data frame (tibble) containing long data.

## Examples

To download all the time series data from an ABS catalogue number to your disk, and import the data to R as a single tidy data frame, use `read_abs()`. Here's an example with the Wage Price Index, catalogue number 6345.0:

```{r all-wpi}
library(readabs)

all_wpi <- read_abs("6345.0")

str(all_wpi)
```

Maybe you only want a particular table? Here's how you get a single table:

```{r wpi1}

wpi_t1 <- read_abs("6345.0", tables = 1)
```

If you want multiple tables, but not the whole catalogue, that's easy too:

```{r wpi1_5}

wpi_t1_t5 <- read_abs("6345.0", tables = c("1", "5a"))

```

For more examples, see the readabs vignette.

## Mentioned in Awesome Official Statistics Software
[![Mentioned in Awesome Official Statistics ](https://awesome.re/mentioned-badge.svg)](http://www.awesomeofficialstatistics.org)

We're pleased to be included in a [list of software](https://github.com/SNStatComp/awesome-official-statistics-software) that can be used to work with official statistics.

## Package history
From version 0.3.0, `readabs` gained new functionality. Pre-0.3.0 functions still work,
but `read_abs_data()` is soft-deprecated and the behaviour of `read_abs_metadata()`
has changed. The old version of `readsabs` is available in the [0.2.9 branch on Github](https://github.com/MattCowgill/readabs/tree/0.2.9).

