---
output: github_document
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

version <- as.vector(read.dcf('DESCRIPTION')[, 'Version'])
version <- gsub('-', '.', version)  
```

# readabs
<!-- badges: start -->
[![Build Status](https://travis-ci.org/MattCowgill/readabs.svg?branch=master)](https://travis-ci.org/MattCowgill/readabs)
[![codecov status](https://img.shields.io/codecov/c/github/mattcowgill/readabs.svg)](https://codecov.io/gh/MattCowgill/readabs)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![CRAN status](https://www.r-pkg.org/badges/version/readabs)](https://cran.r-project.org/package=readabs)
<!-- badges: end -->

## Overview
readabs helps you easily download, import, and tidy time series data from the Australian Bureau of Statistics from within R. This saves you time manually downloading and
tediously tidying time series data and allows you to spend more time on your 
analysis.

We'd welcome Github issues containing error reports or feature requests. 
Alternatively you can email the package maintainer at mattcowgill at gmail dot com.

## Installation

Install the latest CRAN version of **readabs** with:

```{r cran-installation, eval = FALSE}
install.packages("readabs")
```

You can install the developer version of **readabs** from GitHub with:
```{r gh-installation, eval = FALSE}
# if you don't have devtools installed, first run:
# install.packages("devtools")
devtools::install_github("mattcowgill/readabs")
```

## New in recent versions

In version 0.4.0 of the readabs package, 

* A new `separate_series()` function has been added, to be used in conjunction with `read_abs()` to help tidy data (thanks to David Diviny).
* A new `read_cpi()` convenience function returns a tibble containing dates and CPI index numbers.
* The `read_abs()` function now has an optional `series_id` argument, allowing users to get individual time series by using the unique identifiers given to them by the ABS.
* The `read_abs()` function now has a `retain_files` argument (TRUE by default); by setting this to false, downloaded spreadsheets will be stored in a temporary directory
* Some minor bug fixes and enhances - see NEWS for details.

In 0.4.1,

* The `separate_series()` function gained a `remove_nas` argument - thanks to Sam Gow for the suggestion. This makes tidying data easier.
* Various minor improvements and bug fixes.

## Usage

There is one key function in **readabs**. It is:

* `read_abs()`, which downloads, imports, and tidies time series data from the ABS website.

There are some other functions you may find useful.

* `read_abs_local()` imports and tidies time series data from ABS spreadsheets stored on a local drive.
* `separate_series()` splits the `series` column of a tidied ABS time series spreadsheet into multiple columns, reducing the manual wrangling that's needed to work with the data.

Both `read_abs()` and `read_abs_local()` return a single tidy data frame (tibble) containing long data.

## Examples

To download all the time series data from an ABS catalogue number to your disk, and import the data to R as a single tidy data frame, use `read_abs()`. Here's an example with the Wage Price Index, catalogue number 6345.0:

```{r all-wpi}
library(readabs)

all_wpi <- read_abs("6345.0")

str(all_wpi)
```

Maybe you only want a particular table? Here's how you get a single table:

```{r wpi1}

wpi_t1 <- read_abs("6345.0", tables = 1)
```

If you want multiple tables, but not the whole catalogue, that's easy too:

```{r wpi1_5}

wpi_t1_t5 <- read_abs("6345.0", tables = c("1", "5a"))

```

In many cases, the `series` column will contain multiple components, separated by ';'. The `separate_series()` function can help wrangling this column.

For more examples, please see the readabs vignette (run `browseVignettes("readabs")`).

## Mentioned in Awesome Official Statistics Software
[![Mentioned in Awesome Official Statistics ](https://awesome.re/mentioned-badge.svg)](http://www.awesomeofficialstatistics.org)

We're pleased to be included in a [list of software](https://github.com/SNStatComp/awesome-official-statistics-software) that can be used to work with official statistics.

## Package history
From version 0.3.0, `readabs` gained new functionality. Pre-0.3.0 functions still work,
but `read_abs_data()` is soft-deprecated and the behaviour of `read_abs_metadata()`
has changed. The old version of `readabs` is available in the [0.2.9 branch on Github](https://github.com/MattCowgill/readabs/tree/0.2.9).

